---
title: "02_Exercise1"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load additionnal librairies.

By default, only a few packages are loaded in your R workspace. To load additionnal packages, you'll need the function library(). This function works only if the package is already installed. In case the package is not installed, or in case you make a typo in the name of an installed package, you should see an error message like "there is no package called [name-of-your-package]".    

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(here)
library(ggplot2)
library(ggthemes)
```

I recommend to put the librairies you need for your script always at the top of your .Rmd, in a separate chunk.


## Define a theme for the plots

This part is optionnal, but it is recommended to customize the theme of the plots and keep the same theme all along the plots. Below is one proposition of theme.    

```{r, eval=TRUE}
# Define a personnal theme
custom_plot_theme <- function(...){
theme_bw() +
theme(panel.grid = element_blank(),
      axis.line = element_line(size = .7, color = "black"),
axis.text = element_text(size = 11),
axis.title = element_text(size = 12),
legend.text = element_text(size = 11),
legend.title = element_text(size = 12),
legend.key.size = unit(0.4, "cm"),
strip.text.x = element_text(size = 12, colour = "black", angle = 0),
strip.text.y = element_text(size = 12, colour = "black", angle = 90))
}
```


## Import a dataset     

We will now load the "division_timelapse_spot_stats.csv" file (in /dataset/fly_embryo). The here() function ensures to have always a correct path, whatever the location of the R project folder.      

Try typing here() only.
```{r}
here()
```

The read_csv from the library readr read csv files and render them as "tibble", a certain type of dataframe.       
```{r}
df_fly_spots <- read_csv(here("datasets/fly_embryo/division_timelapse_spot_stats.csv"))
```

## View

Let's start by viewing the dataframe. The function head() displays only the 6 first rows.       
```{r, eval=TRUE}
# Preferred method, works with all sizes of dataframes
df_fly_spots %>% 
  head()
```

```{r, eval=FALSE}
# Alternatively, to see the whole table (be careful with big dataframes!!)
df_fly_spots %>% 
  View()
```
df_fly_spots is a tibble (= a certain type of dataframe), with 22 columns and 11 269 rows. On a tibble, we can read the type of the variable in each column below the name of the column: int = integer, chr = character, and dbl = double.    

## Mutate

Let's create a column that contains the results of the computation of the coefficient of variation of the intensity (of fluorescence). The CV is the ratio of the standard deviation over the mean.   

```{r, eval=FALSE, echo=FALSE}
df_fly_spots %>% 
  mutate(MISSING_PART) %>% 
  select(MISSING_PART, MEAN_INTENSITY, STANDARD_DEVIATION) %>% 
  head()
```

```{r, eval=TRUE}
df_fly_spots %>% 
  mutate(CV = STANDARD_DEVIATION/MEAN_INTENSITY) %>% 
  select(CV, MEAN_INTENSITY, STANDARD_DEVIATION) %>% 
  head()
```


NB: to rename a column, we can use rename(). Here we'll rename the column STANDARD_DEVIATION into SD.          
```{r}
df_fly_spots %>% 
  rename(SD = STANDARD_DEVIATION)
```


## Select

We can select only the columns we need with the function select(). For instance, let's select the columns Label, MEAN_INTENSITY, and STANDARD_DEVIATION. NB: in the following chunks, your task is to replace the "MISSING_PART" by the right code. You can find the solutions in the /solutions folder.
```{r, eval=FALSE, echo=FALSE}
df_fly_spots %>% 
  select(MISSING_PART) %>% 
  head()
```

```{r, eval=TRUE}
df_fly_spots %>% 
  select(Label, MEAN_INTENSITY, STANDARD_DEVIATION) %>% 
  head()
```

In practice, it can be interesting to select only a few columns for printing in a report. To display a table in a report, we can use the kable() function from the knitr library.

```{r, eval=FALSE, echo=FALSE}
df_fly_spots %>% 
  select(MISSING_PART) %>% 
  head() %>% 
  knitr::kable()
```

```{r, eval=TRUE}
df_fly_spots %>% 
  select(Label, MEAN_INTENSITY, STANDARD_DEVIATION) %>% 
  head() %>% 
  knitr::kable()
```


We can also remove one or several column with select(). Let's get rid of the column CELL_DIVISION_TIME.      
```{r, eval=FALSE, echo=FALSE}
df_fly_spots <- df_fly_spots %>% 
  select(MISSING_PART)
```

```{r, eval=TRUE}
df_fly_spots <- df_fly_spots %>% 
  select(-CELL_DIVISION_TIME)
```

## Plots  

We would like to see the distribution of the mean fluorescence intensity (per spot) over the whole dataset.     
```{r, eval=FALSE, echo=FALSE}
ggplot(df_fly_spots, aes(x = MISSING_PART)) +
  geom_histogram(bins = 15) +
  custom_plot_theme()
```

```{r, eval=TRUE}
ggplot(df_fly_spots, aes(MEAN_INTENSITY)) +
  geom_histogram(bins = 15) +
  custom_plot_theme()
```

We see roughly a few values around 30000 and all the others are close to 32000.

Let's plot now the mean intensity in function of the time (=FRAME). FRAME is a character variable. It has to be converted to numeric first. We will create a new column called FRAME_num containing the FRAME variable as numeric.       
NB: In the chunk below, the plot was added directly to the chain of commands using the "%>%" (pipe) symbol. All the commands belonging to the plot are still added with a "+".      

```{r, eval=FALSE, echo=FALSE}
df_fly_spots %>% 
  mutate(FRAME_num = as.numeric(MISSING_PART)) %>% 
ggplot(aes(x=FRAME_num, y=MISSING_PART))+
  geom_point() +
  geom_smooth() +
  custom_plot_theme()
```

```{r, eval=TRUE}
df_fly_spots %>% 
  mutate(FRAME_num = as.numeric(FRAME)) %>% 
ggplot(aes(x=FRAME_num, y=MEAN_INTENSITY))+
  geom_point() +
  geom_smooth() +
  custom_plot_theme()
```

With this plot, we don't see much variation of the mean fluorescence intensity over time.

## Filter

We'll now filter the rows and keep only the spots that have a mean fluorescence intensity above 32900 to see if we see better the variations of the mean fluorescence intensity over time.        
          
```{r, eval=FALSE, echo=FALSE}
filtered_df_fly_embryo <- df_fly_spots %>% 
  filter(MEAN_INTENSITY > MISSING_PART) %>% 
  mutate(FRAME_num = as.numeric(MISSING_PART)) 

ggplot(filtered_df_fly_embryo, aes(x=MISSING_PART, y=MISSING_PART))+
  geom_point() +
  geom_jitter() +
  geom_smooth() +
  xlab("Frame") +
  ylab("Mean intensity")
  custom_plot_theme()
```

```{r, eval=TRUE}
filtered_df_fly_embryo <- df_fly_spots %>% 
  filter(MEAN_INTENSITY > 32900) %>% 
  mutate(FRAME_num = as.numeric(FRAME))

ggplot(filtered_df_fly_embryo, aes(x=FRAME_num, y=MEAN_INTENSITY))+
  geom_point() +
  geom_jitter() +
  geom_smooth() +
  xlab("Frame") +
  ylab("Mean intensity") +
  custom_plot_theme()
```
          
We see now some variations around frame 12 to 15. Interestingly, this seems to correspond to a period where most of the cells are dividing in the embryo. In the part below, we'll visualise the number of spots over time.          

## Group by and summarise

We'll count the number of spots by FRAME and store the result in a column called "nb". For this, we need to group the spots by FRAME and count them using the function n().   

```{r, eval=FALSE, echo=FALSE}
nb_nuclei <- df_fly_spots %>% 
  group_by(MISSING_PART) %>% 
  summarise(MISSING_PART = n())
```

```{r, eval=TRUE}
nb_nuclei <- df_fly_spots %>% 
  group_by(FRAME) %>% 
  summarise(nb = n())
```

View the content of nb_nuclei using one of the functions above.    
```{r, eval=FALSE, echo=FALSE}
nb_nuclei %>% 
  MISSING_PART
```

```{r}
nb_nuclei %>% 
  head()
```


## Left join

We would like to add the count of spots stored in nb_nuclei to the main dataframe df_fly_spots. For this, we will use a left_join() and do the join on the column FRAME. We will then display the number of nuclei in function of the FRAME.     

```{r, eval=FALSE, echo=FALSE}
df_fly_spots2 <- left_join(MISSING_PART, nb_nuclei, by = MISSING_PART)

ggplot(MISSING_PART, aes(x = as.numeric(FRAME), y = MISSING_PART)) + 
  geom_point() +
  geom_line() +
  xlab("FRAME") + 
  ylab("Number of nuclei") +
  custom_plot_theme()
```

```{r, eval=TRUE}
df_fly_spots2 <- left_join(df_fly_spots, nb_nuclei, by = "FRAME")

ggplot(df_fly_spots2, aes(x = as.numeric(FRAME), y = nb)) + 
  geom_point() +
  geom_line() +
  xlab("FRAME") + 
  ylab("Number of nuclei") +
  custom_plot_theme()
```


