---
title: "04_Exercise2"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The package you'll explore below is an example of what can be done with R downstream of image analysis softwares like Fiji to build advanced workflows for the analysis of data coming from images.     
If you are working on your own computer, don't forget to install the mamut2r package first (look at Bonus_Install.Rmd).

## Load librairies

As in exercise1, we load first the libraries needed and define a theme for the plots. 

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(here)
library(ggplot2)
library(mamut2r)
library(XML)
library(cellviz3d)
```

```{r, eval=TRUE}
# Define a personnal theme
custom_plot_theme <- function(...){
theme_bw() +
theme(panel.grid = element_blank(),
      axis.line = element_line(size = .7, color = "black"),
axis.text = element_text(size = 11),
axis.title = element_text(size = 12),
legend.text = element_text(size = 11),
legend.title = element_text(size = 12),
legend.key.size = unit(0.4, "cm"),
strip.text.x = element_text(size = 12, colour = "black", angle = 0),
strip.text.y = element_text(size = 12, colour = "black", angle = 90))
}
```


## Open the vignette "Getting started"     

Access the help of the whole mamut2r package and find the vignettes. Open the vignette called "getting started".

```{r, eval=FALSE, echo=FALSE}
help(MISSING_PART)
```


## Import data

With the help of the vignette, find how to import the .xml data ("division_timelapse_trackmate.xml" in datasets/fly_embryo) as two dataframes. Don't forget to use the here() function to specify a path relative to the location of your R project. Importing the .xml might take a few seconds.              
```{r, eval=FALSE, echo=FALSE}
fileXML <- MISSING_PART
Spots_df <- MISSING_PART
Tracks_df <- MISSING_PART
```


## Display the first rows of the dataframes

As in Exercise1, we will check the content of the dataframes prior to any other analysis.      

```{r, eval=FALSE, echo=FALSE}
Spots_df %>% 
  MISSING_PART

Tracks_df %>% 
  MISSING_PART
```


Spots_df and Tracks_df are two dataframes. We can manipulate the data with the library dplyr (select, mutate, filter, group_by, summarise, left_join....) and make plots with the library ggplot2 in the same way as we did for the dataframe in exercise1. Additionnally, some functions of mamut2r allows extra calculations, and vizualisations of the tracks, specific to these data. mamut2r be coupled with another library called cellviz3d for the 3D/4D vizualisation of the spots.           

In this exercise, we would like to explore the possibility to use the variation of fluorescence prior to division to spot the false-detected divisions.      

## Preparation of the data

As we displayed the first rows of the Spots_df dataframe above using the head() function, we have seen that the MEAN_INTENSITY variable was imported as a character (chr). Prior to any further analysis, we'll convert the MEAN_INTENSITY column to numeric.      

```{r, eval=TRUE}
Spots_df <- Spots_df %>% 
  mutate(MEAN_INTENSITY = as.numeric(MEAN_INTENSITY))
```

Do you remember how to calculate the number of spots per frame and add this information into the dataframe?       
```{r, eval=FALSE, echo=FALSE}
nb_nuclei <- MISSING_PART %>% 
  group_by(MISSING_PART) %>% 
  summarise(nb = n())

Spots_df_tmp <- left_join(MISSING_PART, MISSING_PART, by = MISSING_PART)
```


From now on, we will use the function countSpot() from the package mamut2r. It does the same calculation. Functions are pieces of code that have been written in a generic way to be easily re-used several times. Using function shorten and simplify the main code, and ensure the use of the exact same commands. Out of curiosity, you can type countSpot in the console, without the parentheses, to see the code of this function.           
```{r, eval=TRUE}
Spots_df <- countSpot(Spots_df, FRAME)
```


## Visualization in 3D+time of the mean fluorescence intensity   

We will now visualize in 3D+time the mean fluorescence intensity using the function plotlySpots_all() from the package cellviz3d.       

```{r, eval=TRUE}
plotlySpots_all(Spots_df = Spots_df,
                x = POSITION_X,
                y = POSITION_Y,
                z = POSITION_Z,
                timepoint = FRAME,
                number = nb,
                color = MEAN_INTENSITY)
```

As we see here on the 3D vizualisation, and as we saw before on the distribution of mean fluorescence intensities, the spots on the border have a lower fluorescence intensity than the ones in the middle of the embryo. We will filter out the spots with the lowest values (below 32 000, or even below 32 900) to see more clearly the variations over time on the rest of the embryo. How would you do the filtering?      

```{r, eval=FALSE, echo=FALSE}
Spots_df_filtered <- MISSING_PART %>% 
  filter(MISSING_PART)

plotlySpots_all(Spots_df = MISSING_PART,
                x = POSITION_X,
                y = POSITION_Y,
                z = POSITION_Z,
                timepoint = FRAME,
                number = nb,
                color = MISSING_PART)
```


We are now able to visualise the mean fluorescence intensity in 3D+time. As observed in exercise1, variation of the mean fluorescence intensity exist, notably when almost all the cells divide synchronously, with higher fluorescence values prior to division and lower after, but the variations are subtle.             

## Visualization of cells that are about to divide

We are going to look at cells that are about to divide. The Tracks_df contains all information relative to tracks. For each track, and for each frame, it contains the ID of the source spot (SPOT_SOURCE_ID) and of the target(s) (SPOT_TARGET_ID). There can be no target (track is finished), one target (track is continued) or two targets (the track is dividing). The ID of the source spot is repeated in Tracks_df to match the ID(s) of the target(s). We can use the function n() to count the number of occurences of each source, and hence to detect cells that are about to divide. We combine this information to the Spots_df dataframe, using a left_join. How would you do that?        

```{r, eval=FALSE, echo=FALSE}
dividingCell <- Tracks_df %>% 
  group_by(MISSING_PART) %>% 
  summarise(divStatus = MISSING_PART

Spots_df <- left_join(MISSING_PART, MISSING_PART, by = c("ID" = "SPOT_SOURCE_ID"))
```


We will now visualise the outcome of these calculations in 3D+time using the function plotlySpots_all(). Cells that are about to divide have two targets, the others 1, or none if the track is discontinued.          
```{r, eval=TRUE}
Spots_df <- countSpot(Spots_df, FRAME)

plotlySpots_all(Spots_df = Spots_df,
                x = POSITION_X,
                y = POSITION_Y,
                z = POSITION_Z,
                timepoint = FRAME,
                number = nb,
                color = divStatus)
```

## Using the difference of intensity before/after division to find false-detected divisions (bonus)

This last part illustrate how we could try to use the difference of intensity before/after division to find false-detected divisions. 

We now have the information regarding the mean fluorescence intensity and the information regarding the division status (the spot will divide at the next frame or not).      

We'll first calculate the difference of mean intensity before and after division. For this, we'll calculate the difference of mean intensity between consecutive timepoints and group these differences according to the fate of the cell over these two timepoints (the cell divided or not).      
```{r, eval=TRUE}
Spots_df_small <- Spots_df %>% 
  select(ID, MEAN_INTENSITY) %>% 
  filter(MEAN_INTENSITY > 32900)

Tracks_df_diffTargetSource <- Tracks_df %>% 
  left_join(Spots_df_small, by = c("SPOT_SOURCE_ID" = "ID")) %>% 
  rename(MEAN_INTENSITY_SOURCE = MEAN_INTENSITY) %>% 
  left_join(Spots_df_small, by = c("SPOT_TARGET_ID" = "ID")) %>% 
  rename(MEAN_INTENSITY_TARGET = MEAN_INTENSITY) %>% 
  group_by(SPOT_SOURCE_ID) %>% 
  summarise(divStatus = n(),
            meanSOURCE = mean(MEAN_INTENSITY_SOURCE),
            meanTARGET = mean(MEAN_INTENSITY_TARGET)) %>% 
  mutate(diff = meanTARGET - meanSOURCE)
```

The plot below shows the difference of mean fluorescence intensity between two timepoints Tn and Tn+1, for the cells that did not divide (group 1), and for the cells that divided (group 2) between Tn and Tn+1.        
```{r, warning=FALSE}
ggplot(Tracks_df_diffTargetSource, aes(diff)) +
  geom_histogram(bins = 80) +
  geom_vline(xintercept = 0, colour = "red") +
  facet_wrap(facets = vars(divStatus), scales = "free_y", ncol = 1) +
  custom_plot_theme() + 
  xlab("Difference of mean fluorescence intensity")
```


To locate in 3D+time the spots that have a positive difference of mean intensity (e.g. that could be falsely detected), we will add the computed difference to the Spots_df dataframe (containing the x, y, z, t coordinates of the spots) with a left join, and use plotlySpots_all() to display the information.     
```{r, eval=TRUE}
Spots_df2 <- Spots_df %>% 
  left_join(Tracks_df_diffTargetSource, by = c("ID" = "SPOT_SOURCE_ID", "divStatus")) %>% 
  mutate(diff = if_else(divStatus != 2, NA_real_, diff))

Spots_df2 <- countSpot(Spots_df2, FRAME)

plotlySpots_all(Spots_df = Spots_df2,
                x = POSITION_X,
                y = POSITION_Y,
                z = POSITION_Z,
                timepoint = FRAME,
                number = nb,
                color = diff)
```
