---
title: "05_Going_further"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Don't forget to cite R and the packages you use

```{r}
citation(package = "mamut2r")
```
With Zotero (https://www.zotero.org/): copy the BibTeX entry in the clipboard, then in Zotero, go to File -> import from clipboard.

Don't forget to cite R as well.


## Extra ressources about R and image analysis

A webinar on working with images in R: https://vimeo.com/301405941 (R Open Sci community call)         

A few R packages that can be useful in bioimage analysis:       
- momocs: geometric morphometrics (https://vbonhomme.github.io/Momocs/vignettes/Momocs_speed_dating.html)            
- colocr: Co-localization analysis of fluorescence microscopy images  (https://ropensci.org/blog/2018/11/27/colocr/)          
- Rvcg: manipulation of triangular meshes  

## Extra ressources to learn more about R     

An online book: R for Data Science, Garrett Grolemund and Hadley Wickham: https://r4ds.had.co.nz/

A serie of tutorials with videos to get started with R (tidyverse, ggplot2, Rmarkdown...): https://rladiessydney.org/post/2018/11/01/ryouwithme/        

A meetup from the R Ladies Sydney that was live streamed https://rladiessydney.org/post/2018/08/21/little-miss-tidyverse/, the corresponding GitHub repository (that inspired this course) https://github.com/rladiessydney/littlemisstidyverse and slides https://slides.com/djnavarro/tidyverse-for-beginners .

For the french speakers, I also recommend having a look at the blog of Lise Vaudor and her awesome drawings to explain R http://perso.ens-lyon.fr/lise.vaudor/ 


## To get inspired when making plots

https://www.r-graph-gallery.com/      


## Extra ressources about data analysis

Here are two articles from a french company that I translated to english about data formatting and common mistakes in data visualisation:       
https://rtask.thinkr.fr/blog/data-visualisation-pitfalls-how-to-avoid-barbarplots/      
https://rtask.thinkr.fr/blog/the-ten-commandments-for-a-well-formatted-database/       


## Extra ressources to get help

https://stackoverflow.com/
http://www.sthda.com/english/     
All the R studio cheatsheets https://www.rstudio.com/resources/cheatsheets/   

How to get help: https://masalmon.eu/2018/07/22/wheretogethelp/          

The reprex R package, to produce reproducible example ("reprex") that you can share with others: https://reprex.tidyverse.org/      


## To keep informed on what is going on in the R community

This websites aggregate content from other blogs about R:       
https://www.r-bloggers.com/        
https://rweekly.org/        

On Twitter, you can have a look at #rstats   


## To join local or remote R groups

UseR meetups: https://www.meetup.com             
Rladies groups: https://rladies.org/                
R ladies remote group (sign in form): https://twitter.us18.list-manage.com/subscribe?u=2ea47051e532678beaba00ee9&id=0d53e23281    
French speaking R remote group "Grrr". To sign in: https://frama.link/r-grrr          
