---
title: "Bonus_Install"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to install the packages we will use for the course

This part is just a reminder on how to install the packages on your own computer. For the training school 10 (TS10), the packages are already installed.

## Installing the tidyverse packages 

To install a package, the command is: 
```{r, eval=FALSE}
install.packages("NAME_of_the_package")
```
Or from the Rstudio menu: `Tools -> Install packages...`      

For instance:        
```{r, eval=FALSE}
install.packages("dplyr")
```

N.B.: I set the eval preference of the chunks to FALSE to prevent accidental execution of the chunks if you knit this .Rmd. Installation of a package can take a couple of minutes.              

## Installing the mamut2r package

The package I developped is not available on CRAN yet and hence cannot be installed using the install.packages() function. This package is on GitHub. As written in the readme of my package (https://github.com/marionlouveaux/mamut2r), the command to install it is:       

```{r, eval=FALSE}
devtools::install_github("marionlouveaux/mamut2r")
```

You can set `build_vignettes = TRUE` to knit the vignettes.        



